---
title: "Template"
author: "Your Name"
date: 'YYYY-MM-DD (updated: `r Sys.Date()`) '
output:
  word_document:
    toc: yes
    toc_depth: 2
urlcolor: blue
---

```{r setup, include=FALSE}
# By default, all codes are hidden.
knitr::opts_chunk$set(message = FALSE, warning = FALSE, include = FALSE, comment = NA, ft_max_row = 6, tab.topcaption = FALSE, ft.align = "left")
library(knitr)
library(tidyverse)
# remotes::install_github("Bolin-Wu/workshopr", subdir = "rpackage", force = TRUE)
library(workshopr)
library(flextable)
set_flextable_defaults(font.size = 11, padding = 3)
```

# Introduction

This document is to demonstrate some rmd basic functions.


# Flextable
`flextable` is a useful package for printing tables for **word** file output. It is similar to `kable()` for PDF file output.


If one wants to generate tables by loop:

```{r, include=TRUE, echo=FALSE, results='asis'}
for (i in 1:10) {
  cat("Table: ", i)
  ft <- flextable(cars[i, ])
  caption_name <- paste0("Print row: ", i)
  ft <- set_caption(ft, caption_name)
  ft <- autofit(ft)
  flextable_to_rmd(ft)
}
```

This could be useful when one needs to do repetitive work, e.g. printing significant variables from `lm()` function. Please note that the result has to be set to 'asis'.


# Visualization

When you compile the rmd, plots can be generated automatically.

```{r, include = TRUE, eval=TRUE, echo=TRUE}
cbPalette <- c(
  "#A50026", "#D73027", "#F46D43", "#FDAE61", "#FEE090", "#FFFFBF",
  "#E0F3F8", "#ABD9E9", "#74ADD1", "#4575B4", "#313695"
)
fake_data %>%
  select(Lopnr, contains("mmse")) %>%
  pivot_longer(
    cols = contains("mmse"),
    names_to = "wave", values_to = "mmse",
    names_prefix = "mmse_"
  ) %>%
  ggplot(aes(x = wave, y = mmse, fill = wave)) +
  geom_boxplot(na.rm = T) +
  scale_fill_manual(
    name = "MMSE Measures",
    values = cbPalette
  )
```

# Descriptive statistics

Just use `describe` from `psych` as an example. 

```{r , include = TRUE, eval=TRUE, echo=TRUE}
# install.packages('psych')
library(psych)

kable(describe(fake_data, na.rm = TRUE), 
      format='markdown',
      caption="Example",
      digits=2)
```

