---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# workshopr

<!-- badges: start -->
<!-- badges: end -->

The goal of `workshopr` is to make example data and script easily accessible.

# Installation

You can install the development version of workshopr from [GitHub](https://github.com/Bolin-Wu/workshopr):

``` r
install.packages("remotes")
remotes::install_github("Bolin-Wu/workshopr", subdir = "rpackage", force = TRUE)
```

# Data

Author: [Xin Xia](https://ki-su-arc.se/staff/xin-xia/).

```{r example_data, eval=FALSE}
workshopr::fake_data 
```

# R script

Please see below to retrieve code used in this workshop.

## Loop 
Author: [Ashley Tate](https://staff.ki.se/people/ashley-tate).

Load the example code:
```
workshopr::get_code_2023(session = "loop")
```

## Tidyverse & Rmarkdown
Author: [Bolin Wu](https://staff.ki.se/people/bolin-wu).

Load example R code:
```
workshopr::get_code_2023(session = "tidyverse")
```

Get the rmd templates with html and word output:
```
workshopr::get_rmd_2023(name = "pdf_example",output_file = "pdf")
workshopr::get_rmd_2023(name = "word_example",output_file = "word")
workshopr::get_rmd_2023(name = "html_example",output_file = "html")
```
# Issues

You can raise bug reports [here](https://github.com/Bolin-Wu/workshopr/issues).
