---
title: "Final exercise"
author:
  - name: "Xin Xia"
    affiliation: "KI"
date: '2024-04-12 (updated: `r Sys.Date()`) '
output:
  pdf_document: 
    toc: true
    toc_depth: 2
urlcolor: blue
  # rmdformats::downcute:
  #   self_contained: true
  #   thumbnails: true
  #   lightbox: true
  #   gallery: false
  #   highlight: tango
  #   df_print: paged
  #   
    
---

```{r setup, include=FALSE, purl=FALSE}
# Default chunk settings
knitr::opts_chunk$set(message = FALSE, warning = FALSE, include = TRUE)
```

```{r}
# run the install command if did not do it before
# install.packages('remotes')
# remotes::install_github("Bolin-Wu/workshopr", subdir = "rpackage", force = TRUE)

# load the package
library(workshopr)
library(tidyverse)
library(here)
```

# Introduction

Generate a html/word report based on the simulated data set “fake_data” using (1) the rmarkdown templates created by Bolin Wu, (2) loop/apply functions, and (3) rmarkdown cheatsheet (https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf). Feel free to build on the syntaxes Bolin Wu and Ashley Tate wrote. 

# Optional tasks
**The report can include the following contents:**

* Summary statistics of the datasets in a table (mean and standard deviation for continuous variables, count and percentage for categorical variables, the number of missing values in each variable). 

*Hints: use “mutate” (from tidyverse) to define variable classes; use “select(where())” (from tidyverse) to select variables; use “summarise()” (from tidyverse) or “sapply()” to summarise.*

* Summarise the associations of MMSE score at wave 1 with age at wave 1, sex, and educational levels in a table.

*Hints: use “tapply()” to summarise across sex strata; write loops to summarise multiple variables.*

* Make any graph you want to make with the simulated data. If you don’t know what to draw, try to make a spaghetti plot of MMSE changes over time.

*Hint: use “pivot_longer()” to reshape data from a wide format to a long format.*

* In a table, summarise the results of a regression. If you don’t know which model to build, try summarising the output of this model: glm(formula = mmse_wave1~age_base+sex+education, family = "gaussian",data = fake_data).

*Hints: first, check the elements in the output of the model summary by typing summary(model_name)$, confint(model_name), etc; build an empty data frame with names of quantities (e.g., coefficients); use loops and “rbind” to add model results into the empty data frame.*

# Alternatives

* Using the simulated dataset, create a report and decide the contents of the report on your own.
* Create a report using your own data.

# Solutions

You can find solutions from Xin Xia by

```{r, eval=FALSE}
workshopr::get_solution_2023(name = "practice_solution")
```

